3D Regression Heat Map Analysis of Population Study Data

### 2015-06-23 draft.tex
revised figures to save space, removed workflow figure

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{3D Regression Heat Maps}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare 3D Regression Heat Maps.
 %%
 [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{3D Regression Heat Map} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

\textbf{Details on demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
%%
\paragraph{Hypothesis-free and hypothesis-based analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches of analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis allows to derive new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into the \emph{3D Regression Heat Map} prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 %%
 (b) 3D heat map visualization, showing values above the matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D heat map visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%
 }
  \label{fig:System}
\end{figure}
%%
%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
%%
\noindent We designed our system to be openly accessible and easy to use.
%%
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses with other (statistical) tools.
%%
\magenta{This is, however, out of the scope of this work.}
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.

>>>>

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
% \begin{figure}[htb]
%  \centering
%  \includegraphics[width=1.0\linewidth]{figures/workflow}
%  \caption{
%  Different workflow types using the \emph{3D Regression Heat Map}.
%  %%
%  [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
%  %%
%  [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
%  %%
%  The user has the option to either adjust the formula, adjust the transfer function or to derive details on demand on models.
%  %%
%  [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
%  %%
%  The latter is used to compare \emph{3D Regression Heat Maps}.
%  %%
%  [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
%  }
%   \label{fig:Workflow}
% \end{figure}
%%
\noindent Our \emph{3D Regression Heat Map} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

\textbf{Details on demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
%As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
The workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
%%
\paragraph{Hypothesis-free and hypothesis-based analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches of analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis allows to derive new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into our prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature and calculation rules.
 %%
 (b) 3D heat map visualization, showing values above the matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D heat map visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%
 }
  \label{fig:System}
\end{figure}
%%
%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
%%
\noindent We designed our system to be openly accessible and easy to use.
%%
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses with other (statistical) tools.
%%
\magenta{This is, however, out of the scope of this work.}
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
